---
title: "Lists - In Class Exercises"
author: "Dr. Shoemaker"
date: "DATA 2401"
output:nhtml_document

---

# Exercises: 

Make a list that contains the information on a single playing card. Suit, face, and value. 

```{r}
playing <- list("Diamonds", "Spades", "Clubs", "Hearts", "Ace", 2:10, "Jack", "Queen", "King", 1:13)
playing
```


Make a second list where each element contains all the options for suit, face, and value, respectively. 

```{r}
playing_card <- list(
  suit=c("Diamonds", "Spades", "Clubs", "Hearts"),
  face=c("Ace", 2:10, "Jack", "Queen", "King"),
  value=c(1:13)
)
playing_card

```


If `Newlist <- list(a=1:10, b="Good morning", c="Hi")`, write an R statement that will add 1 to each element of the first vector in Newlist.

```{r}
Newlist <- list(a=1:10, b="Good morning", c="Hi")
Newlist[[1]] <- Newlist$a + 1
Newlist
```

If `b <- list(a=1:10, c="Hello", d="AA")`, write an R expression that will give all elements of the first vector of b, except the second, and write another that returns only elements of `a` that are greater than 7.

```{r}
b <- list(a=1:10, c="Hello", d="AA")
b$a[-2]
b$a[8:10]

```

Let `x <- list(a=5:10, c="Hello", d="AA")`, write an R statement to add a new item z = "NewItem" to the list x.

```{r}
x <- list(a=5:10, c="Hello", d="AA")
x[["z"]] <- "NewItem"
x

```


```{r personal-practice}
# New list methods
x <- c("First"=1/2, "Second"=1/4, "Third"=1/4, "4"=3, "5"=0)
x
x+1
x[c(-1,-2,-3)] # "-" means "not", "c(-2:-4)" means "not 2 through 4 (inclusive)"
x["Second"]

# names function
y <- 1:4
names(y) <- c("a", "b", "three", "four") # adds names as index to an existing list
y
y["a"]
names(y) # returns names of indices
names(y)[4] <- "new four"
names(y)

# %in% , xor , any , all logical operators
numlist <- 1:8
8 %in% numlist # is 8 in numlist? T/F
9 %in% numlist
xor(1 %in% numlist, 9 %in% numlist) # T bc only 1 is true
any(4 %in% numlist, 10%in%numlist) # T bc at least 1 is true
all(1%in%numlist, 2%in%numlist, 3%in%numlist, 100%in%numlist) # F bc not all true

# other logical operators (> < >= <= == !=)
1%in%numlist & 2%in%numlist # T bc both true
1%in%numlist | 2%in%numlist # T bc at least one true
!1%in%numlist # F bc condition is true

# Making basic lists
address_info <- list(
  street="23011 Lake Crest Dr",
  city="Huffman",
  state="Texas",
  zip="77336"
)
address_info

family_info <- list(
  names=c("Rex", "Shonna", "Ethan"),
  relativity=c("Father", "Mother", "Brother"),
  ages=c(55, 49, 16),
  is_old=c(TRUE,TRUE,FALSE)
)
family_info
str(family_info)
# note, tags (labels in the list) aren't necessary

# Dollar sign $ notation
family_info$relativity
  # if you have a list within a list, you can use 2 or more $
  # ie my_dogs$favorites$toy produces the list of toys within the 
  # favorites list within the my_dogs list

# Double bracket notation [[]]
family_info[[3]] # gives the 3rd component of the list without label
family_info[3] # gives the 3rd component of the list with label
# my_dogs$favorites[[2]] gives the 2nd component of the list inside the list

# Matrices
numbers <- c(5, 5, 5, 6, 6, 6, 7, 7, 7)
num_matrix <- matrix(numbers, nrow=3)
num_matrix
```

